---
title: "table"
author: "TJ Tang"
date: "4/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
stats_simdata = function(idx){
  data_path = paste0("data/",idx,'.csv')
  sim_data = read.csv(data_path)
  fs = sim_data$full_sample
  ss = sim_data$sample_split
  bfs = (fs - 1) 
  bss = (ss - 1) 
  return(c(mean(bfs,na.rm=T),mean(bss,na.rm=T),sd(bfs,na.rm = T),sd(bss,na.rm = T)))
}
```

K=2,5,full, xgboost, lasso

Table 1: normal, linear, var=0.1, n=50, ncov=5

k = 2, xgboost,  normal, linear, var=0.1, n=50, ncov=5 idx = 1
k = 5, xgboost,  normal, linear, var=0.1, n=50, ncov=5 idx = 2
k = 2, lasso,  normal, linear, var=0.1, n=50, ncov=5 idx = 65
k = 5, lasso,  normal, linear, var=0.1, n=50, ncov=5 idx = 66


Table 1: normal, linear, var=0.1, n=50, ncov=5

```{r}
idx = c(1,2,65,66)
stats = c()
for(i in idx){
  stats = c(stats,stats_simdata(i))
}
stats
```
Table 2: normal, linear, var=0.1, n=50, ncov=100

k = 2, xgboost,  normal, linear, var=0.1, n=50, ncov=100 idx = 3
k = 5, xgboost,  normal, linear, var=0.1, n=50, ncov=100 idx = 4
k = 2, lasso,  normal, linear, var=0.1, n=50, ncov=100 idx = 67
k = 5, lasso,  normal, linear, var=0.1, n=50, ncov=100 idx = 68

